---
title: "Lab 1: Lattice package"
author: "Makesh Srinivasan"
date: "1/22/2022"
output: html_document
---

<h3>Data visualisation</h3><br>
Registration number: 19BCE1717<br>
Faculty: Prof. Parvathi R<br>
Slot: L55 + L56<br>
Course code: CSE3020 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## Instructions:
Explore the lattice package in R using the HSB2 dataset and the builtin MTCars dataset. Run the code given in class <br>


## Sections:
* <a href="#part1">Part 1: HSB2 dataset</a>
* <a href="#part2">Part 2: MTCars dataset</a>

***

<h2 id="part1">PART 1: HSB2 dataset</h2>

##### Import required package and dataset:
```{r}
library(lattice)
df = read.csv("hsb2.csv")
head(df, 5)
```

##### Create factors with value labels
```{r}
ses.f <-factor(df$ses,levels=c(1,2,3),
labels=c("ses_1","ses_2","ses_3"))

prog.f <-factor(df$prog,levels=c(1,2,3),
labels=c("prog_1","prog_2","prog_3"))
```

##### xy Scatter plot
```{r}
xyplot(df$read~df$write)
```

##### Scatterplots for each combination of two factors
```{r}
xyplot(df$read~df$write|ses.f*prog.f,
main="Scatterplots by Cylinders and Gears",
ylab="Miles per Gallon", xlab="Car Weight")
```

##### XYplot
```{r}
xyplot(read~write,
       data = df,
       type = c("p", "r"),
       main = "Relation between Read and Write",
       xlab = "Read Score",
       ylab = "Write Score")
```


##### Multivariate xy scatter plot with customizations

```{r}
xyplot(df$read~df$write | ses.f,
       data = df,
       type = c("p", "r"),
       groups = ses.f,
       main = "Relation between Read and Write",
       xlab = "Read Score",
       ylab = "Write Score")
```


##### Kernel density plot
```{r}
densityplot(~df$read,
main="Density Plot",
xlab="Read score")
```

##### Kernel density plots by factor level
```{r}
densityplot(~df$read|ses.f, data = df,
main="Read Score by SES",
xlab="Read Score")
```



##### Kernel density plots by factor level (alternate layout)
```{r}
densityplot(~df$read|ses.f,
main="Density Plot by SES",
xlab="Read score",
layout=c(1,3))
```

##### Kernel Density Plot together for all factors without points
```{r}
densityplot(~df$read,
groups = ses.f,
plot.points = FALSE,
main = "Kernel density plot over SES",
xlab = "Read score")
```

##### Boxplot
```{r}
bwplot(prog.f ~ df$read | ses.f,
xlab = "SES",
ylab = "Prog",
Main = "Prog given by SES")
```


##### Boxplots for each combination of two factors
```{r}
bwplot(ses.f~df$read|prog.f,
ylab="SES", xlab="PROG",
main ="SES and Prog")
```


##### Boxplot associated with multiple variables and alternate layout
```{r}
bwplot(prog.f ~ df$read |ses.f,
xlab = "READ score",
ylab = "PROG",
Main = "SES and read score",
layout = c(1, 3))
```


##### 3D plot
```{r}
# 3 d plot
cloud(df$socst~df$science*df$math,main = "3D scatterplot")
```


##### 3d scatterplot by factor level
```{r}
cloud(df$science~df$math*df$socst|ses.f,
main="3D Scatterplot by SES")
```

##### Dotplot for each combination of two factors
```{r}
dotplot(ses.f~df$read|prog.f,
main="Dotplot Plot by SES and prog",
xlab="Read score")
```


##### Scatterplot matrix
```{r}
splom(df[c(1,3,4,5,6)],
main="hsb2 Data")
bwplot(prog.f ~ df$read | ses.f,
data = df,
xlab = "Read score",
ylab = "PROG",
Main = "PROG and SES",
panel = panel.violin)
```

##### Contour plot
```{r}
data <- dimnames(volcano)
contour(x=volcano, xlab = "Row", ylab = "Column") # Contour plot for volcano dataset
filled.contour(volcano)
filled.contour(volcano,color.palette = terrain.colors)
```

***

<center><h3>The same is repeated using the MTCars dataset below</h3></center><br>

***

<h2 id="part2">PART 2: MTCars datset</h2>
##### Import necessary package and datset:
```{r}
library(lattice) # already imported from part 1
attach(mtcars)
mtcars
```

##### Create factors with value labels
```{r}
gear.f<-factor(gear,levels=c(3,4,5),
labels=c("3gears","4gears","5gears"))
cyl.f <-factor(cyl,levels=c(4,6,8),
labels=c("4cyl","6cyl","8cyl"))
```

##### xy Scatter plot
```{r}
xyplot(mpg~wt)
```

##### Scatterplots for each combination of two factors
```{r}
xyplot(mpg~wt|cyl.f*gear.f,
main="Scatterplots by Cylinders and Gears",
ylab="Miles per Gallon", xlab="Car Weight")
```

##### XYplot
```{r}
#XYplot 1
xyplot(mpg~wt,
data = mtcars,
type = c("p", "r"),
main = "Relation between wt and mpg",
xlab = "Weight in lbs",
ylab = "Miles/Gallon (US)")

#XYplot 2
xyplot(mpg~wt | cyl.f,
type = c("p", "r"),
groups = cyl.f,
main = "Relation between wt and mpg over cylinders",
xlab = "Weight in lbs",
ylab = "Miles/Gallon (US)")
```

##### Multivariate xy scatter plot with customizations
```{r}
xyplot(mpg~wt | cyl.f,
type = c("p", "r"),
groups = cyl.f,
main = "Relation between wt and mpg over cylinders",
xlab = "Weight in lbs",
ylab = "Miles/Gallon (US)")
```

##### Kernel density plot
```{r}
densityplot(~mpg,
main="Density Plot",
xlab="Miles per Gallon")
```

##### Kernel density plots by factor level
```{r}
densityplot(~mpg|cyl.f,
main="Density Plot by Number of Cylinders",
xlab="Miles per Gallon")
```

##### Kernel density plots by factor level (alternate layout)
```{r}
densityplot(~mpg|cyl.f,
main="Density Plot by Numer of Cylinders",
xlab="Miles per Gallon",
layout=c(1,3))
```

##### Kernel Density Plot together for all factors without points
```{r}
densityplot(~mpg,
groups = gear.f,
plot.points = FALSE,
main = "Kernel density plot over number of gears",
xlab = "Miles Per Gallon (US)")
```

##### Boxplot
```{r}
bwplot(gear.f ~ mpg | cyl.f,
xlab = "Miles per Gallon (US)",
ylab = "No of Gears",
Main = "Mileage by no. of gears and cylinders")
```

##### Boxplots for each combination of two factors
```{r}
bwplot(cyl.f~mpg|gear.f,
ylab="Cylinders", xlab="Miles per Gallon",
main ="Mileage by Cylinders and Gears")
```

##### Boxplot associated with multiple variables and alternate layout
```{r}
bwplot(gear.f ~ mpg |cyl.f,
xlab = "Miles per Gallon (US)",
ylab = "No of Gears",
Main = "Mileage by no. of gears and cylinders",
layout = c(1, 3))
```

##### 3D plot
```{r}
# 3 d plot
cloud(mpg~wt*qsec,main = "3D scatterplot")
```

##### 3d scatterplot by factor level
```{r}
cloud(mpg~wt*qsec|cyl.f,
main="3D Scatterplot by Cylinders")
```

##### Dotplot for each combination of two factors
```{r}
dotplot(cyl.f~mpg|gear.f,
main="Dotplot Plot by Number of Gears and Cylinders",
xlab="Miles Per Gallon")
```

##### Scatterplot matrix
```{r}
splom(mtcars[c(1,3,4,5,6)],
main="MTCARS Data")
bwplot(gear.f ~ mpg | cyl.f,
data = mtcars,
xlab = "Miles per Gallon (US)",
ylab = "No of Gears",
Main = "Mileage by no. of gears and cylinders",
panel = panel.violin)
```

##### Contour plot
```{r}
data <- dimnames(volcano)
contour(x=volcano, xlab = "Row", ylab = "Column") # Contour plot for volcano dataset
filled.contour(volcano)
filled.contour(volcano,color.palette = terrain.colors)
```

***


