---
title: "Lab 1: Introduction to R"
author: "Makesh Srinivasan"
date: "1/7/2022"
output: html_document
---
<center><h1>Data visualisation</h1></center><br>
Registration number: 19BCE1717<br>
Faculty: Prof. Pavithra R<br>
Slot: L55 + L56<br>
Course code: CSE3020 <br>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*** 
## Part 1: Basic Datatypes in R
```{r}
rm(list=ls())
#1) Assign a decimal value to a variable A and display it.
A <- 10.5
print(A)
#2) Print the class name of the variable A.
print(class(A))
#3) Check whether the variable A is of type 'numeric'.
print(is.numeric(A))
#4) Assign an integer value to a variable B and display it.
B <- 5L
print(B)
#5) Check whether the variable B is of type 'integer'.
print(is.integer(B))
#6) Create a variable C which stores the integer part of variable A.
C=as.integer(A)
print(C)
#7) Compute the cost of a chocolate.
cost <- C/B
print(cost)
#8) Represent the money as character string.
as.character(A)
#9) Store the first name and last name of the kid.
fname="Ram"
lname="kumar"
print(is.integer(fname))
#10) Display the message "Ram bought <B> chocolates"
print(sprintf("%s bought %d chocolates", paste(fname,lname), B))
#11) Extract the substring "Little" from the rhymes.
rhymes="Twinkle Twinkle Little Star"
print(substr(rhymes,start=17,stop=22))
#12) Replace "Little" as Big.
print(sub("Little", "Big", rhymes))
#13) Assign a complex number to a variable X.
X <- 2+4i
X
#14) Display the real part of X.
print(Re(X))
#15) Display the imaginary part of X.
print(Im(X))
#16) Compute square root of a negative number
print(sqrt(as.complex(-2))) #sqrt(-2+0i) also works
#17) Check whether real part of X is greater than its imaginary part.
chk <- Re(X)>Im(X)
print(chk)
```

***

## Part 2: Working with vectors
```{r}
rm(list=ls())
#1. Create vector 'class' to store the class names 'class1','class2',.,'class5'
class=c("class1", "class2","class3","class4","class5")
print(class)
#2. Use assign() function to create a vector 'avg' to store the average marks.
assign("avg",c(63.5,72.3,88.9,65.4,79.8))
print(avg)
#3. Display the average mark of class2.
print(avg[2])
#4. Combine the vectors 'class' and 'avg' as details.
details=c(class,avg)
print(details)
#5. Find the length of combined vector 'details'.
print(length(details))
#6. Find the minimum average mark and print the class which scored it.
print(min(avg))
print(class[which.min(avg)]) #which.min(avg)returns the index of the minimum value in vector avg
#7. Find the maximum average mark and print the class which scored it.
print(max(avg))
print(class[which.max(avg)])
#8. Find the total of average marks scored by all classes.
print(sum(avg))
#9. Find the mean of the average marks scored by all classes.
print(mean(avg))
#10. Find the standard deviation of the average marks scored by all classes.
print(sd(avg))
#11. Arrange the average marks in ascending order.
print(sort(avg))
#12. Create a vector classes by repeat the vector class twice.
classes <- rep(class,times=2)
print(classes)
#13. Create a vector marks by repeating each average mark twice.
marks <- rep(avg,each=2)
print(marks)
#14. Create a sequence of 10 to 1. Add it to the vector avg and display it.
s=seq(10,1)
print(avg+s)
#15. Create a vector bool that contains logical values 'TRUE' or 'FALSE' depending on the condition average marks>70.
bool=avg>70
print(class[bool])
```

***

## Part 3: Working with matrix
```{r}
rm(list=ls())
#1. Represent the height information of a team of 12 basketball players as a matrix of dimension 4x3 in row major form.
height=matrix(c(162,173,160,181,159,176,183,162,181,160,170,172),nrow=4,ncol=3,byrow=TRUE)
print(height)
#2. Display the height at row 3 and column 2.
print(height[3,2])
#3. Display all the heights in row 2.
print(height[2,])
#Display all the heights in column 3.
print(height[,3])
#5. Extract the heights in all rows but only in column 1 and 3.
print(height[,c(1,3)])
#6. Find the transpose of the matrix.
print(t(height))
#7. Four more players got added to the team. Update the matrix to reflect the heights of the players.
height=cbind(height,c(176,168,161,172))
print(height)
#8. Append four more players' height in the matrix.
height=rbind(height,c(175,162,170,165))
print(height)

```

***

## Part 4: R commands from the PPT

***

##### Basic mathematical operations:
```{r}
# Root
print(sqrt(2))

# Log
print(log(2))

# Assignment and arithmetic (+) operation
x = 5
y = 10
z <- x + y
print(z)

# floor:
floor(5.555)

# round:
round(5.555)
```

##### Generate a vector with common difference:
```{r}
v = seq(1,5, by=.5)
v
```

##### Vector operations:
```{r}
v1 = c(6,5,4,3,2,1)
v2 = c(10,9,8,7,6,5)

# Arithmetic operation on the vectors
v3 = v1 + v2
print(v3)

vector = c(5,1,46,5,4,3,2,1)
sorted_vector <- sort(vector)

# Sort
sorted_vector

# Rank
rank(vector)

# Order
order(vector)

# Which
which(vector == 16)
```

##### Operations of V3 vector:
```{r}
# I have also used the paste and print functions here
print("Operations of V3 vector")
print(paste("Maximum = ", max(v3)))
print(paste("Minimum = ", min(v3)))
print(paste("Mean = ", mean(v3)))
print(paste("Standard deviation = ", sd(v3)))
print(paste("Length of the vector V3 = ", length(v3)))
```

##### Sampler:
```{r}
#Without replacement
sample(1:40, 5)

#With replacement
sample(1:50, 10, replace=T)
```

##### R bind:
```{r}
a <- c(1, 2, 3, 4, 5)
b <- c(6, 7, 8, 9, 10)
r_matrix <- rbind(a, b)
r_matrix
```

##### C Bind:
```{r}
c_matrix <- cbind(a, b)
c_matrix
```

##### Repetition to create a vector:
```{r}
rep(c(1:4), 3)
```

##### Random number generator with levels:
```{r}
gl(2, 10, length=20)
```

<center><h3>Dataframe</h3></center>

##### Inbuilt: Import and load the Iris dataset from the MASS package:

```{r}
library("MASS")
data(iris)
head(iris, 5)
```

##### User generated dataframe:

```{r}
students <- c("Makesh", "Anthra", "Bettina", "Chandhru", "Shyam")
event_1 <- c(10,10,6,8,10)
event_2 <- c(5,7,9,0,0)
event_3 <- c(7,9,0,8,6)
event_4 <- c(9,7,8,10,10)
data <- data.frame(students, event_1, event_2, event_3, event_4)
data
```

##### Read from the directory:
```{r}
directory_read <- read.csv("sample.csv")
head(directory_read)
```

##### Explore the dataset using common functions

Dimensions:
```{r}
dim(iris)
```

Structure:
```{r}
str(iris)
```

Use table() function:
```{r}
table(iris$Sepal.Length)
```

Names and attributes:
```{r}
names(iris)
attributes(iris)
```

Tail and head (3) tuples:
```{r}
head(iris, 3)
tail(iris, 3)
```

Summary:
```{r}
summary(iris)
```

Quantile:
```{r}
quantile(iris$Sepal.Length)
quantile(iris$Sepal.Length, c(.1, .3, .65)) # choose the quantiles
```

Variance:
```{r}
var(iris$Sepal.Length)
```

Covariance:
```{r}
cov(iris$Sepal.Length, iris$Petal.Length)
```

Covariance matrix:
```{r}
cov(iris[,1:4])

```

Correlation and correlation matrix:
```{r}
cor(iris$Sepal.Length, iris$Petal.Length)
cor(iris[,1:4])
```

Aggregate:
```{r}
aggregate(Sepal.Length ~ Species, summary, data=iris)
```


##### apply():
```{r}
m <- matrix(C<-(1:10),nrow=5, ncol=5)
#Matrix:
m

#Apply()
apply(m, 2, sum)
```

##### lapply():
```{r}
names <- c("NAME1","NAME2","NAME3","NAME4")
names_lower <-lapply(names, tolower)
str(names_lower)
```

##### sapply():
```{r}
#same as lapply but a vector is returned
smn <- sapply(iris$Sepal.Length, min)
smn
```

***

<center><h3>Probability distribution</h3></center>

##### Binomial distribution:
```{r}
dbinom(2, 3, 0.60)
```

##### Poisson distribution:
```{r}
dpois(2, 1)
```

##### Normal probability distribution:
```{r}
pnorm(150, 156, 4.6)
```

***

<center><h3>Plots</h3></center>

##### Histogram:
```{r}
hist(iris$Sepal.Length)
```
##### Density plot:
```{r}
plot(density(iris$Sepal.Length))
```
##### Pie chart:
```{r}
pie(table(iris$Species)) 
```
##### Barplot:
```{r}
barplot(table(iris$Species))
```
##### Boxplot:
Individual:
```{r}
boxplot(iris$Sepal.Length)
```

Grouped:
```{r}
boxplot(iris$Sepal.Length,iris$Sepal.Width)
```

```{r}
boxplot(Sepal.Length~Species, data=iris)
```
##### Species-wise scatter plot: 
```{r}
with(iris, plot(Sepal.Length, Sepal.Width, col=Species, pch=as.numeric(Species)))
```
##### Pair plot:
```{r}
pairs(iris)
```

##### 3D plots:
```{r}
# Import a library to view 3d plots
library("scatterplot3d")

scatterplot3d(iris$Petal.Width, iris$Sepal.Length, iris$Sepal.Width)
```
##### Heatmap
```{r}
distMatrix <- as.matrix(dist(iris[,1:4]))
heatmap(distMatrix)
```
##### Contour plot:
```{r}
filled.contour(volcano, color=terrain.colors, asp=1,plot.axes=contour(volcano, add=T))
```
##### Parallel coordinates:
```{r}
parcoord(iris[1:4], col=iris$Species)
```
##### Parallel plots:
```{r}
library(lattice)
parallelplot(~iris[1:4] | Species, data=iris)
```

##### Line plot:
```{r}
plot(iris$Sepal.Length, iris$Sepal.Width, type="l") 
```

##### Panelling graphs:
```{r}
par(mfrow=c(2,2))
hist(iris$Sepal.Length, main='Histogram',xlab='Sepal length', ylab ='Frequency', col=heat.colors(14))

boxplot(iris$Sepal.Length, iris$Sepal.Width, iris$Petal.Length, iris$Petal.Width,  main='Iris boxplot of features', ylab='f',xlab='Shift', names = c('mm','mm','mm','mm'))

plot(iris$Sepal.Length,iris$Sepal.Width,main='Sepal length vs width', xlab='mm',ylab='mm',pch=2)
plot(iris$Petal.Length,iris$Sepal.Length,type="l",main='petal vs sepal length',xlab='mm',ylab='mm')
lines(iris$Petal.Length,iris$Sepal.Length,lty=2)

```

***

<center><h3>Loops, functions and conditionals</h3></center>

##### Functions, loops and conditionals:
```{r}
# I made a palindrome checker to perform all three things mentioned above.

isplaindrome <- function(n){
    reverse = 0
    buff = n
    while (n > 0) {
      r = n %% 10
      reverse = reverse * 10 + r
      n = n %/% 10
    }
    if (reverse == buff)
    {
      print(paste("It is palindrome!", reverse))
    }
    else{
      print(paste("Not a palindrome", reverse))
    }
}

## Call the function:
isplaindrome(12345) # should return "Not a palindrome"
isplaindrome(123321) # should return "It is palindrome!"
```

<br>
<br>
NOTE: The clustering parts are not yet covered in class and will be presented in the later lab exercises

***

