---
title: "Lab 2: ggplot2 package"
author: "Makesh Srinivasan"
date: "1/28/2022"
output: html_document
---

<h3>Data visualisation</h3><br>
Registration number: 19BCE1717<br>
Faculty: Prof. Parvathi R<br>
Slot: L55 + L56<br>
Course code: CSE3020 <br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## Instructions:
Explore the ggplot2 package in R using a dataset of your choice and the midwest dataset used in class. Run the code given in class <br>


## Sections:
* <a href="#part1">Part 1: Survey dataset (data of your choice)</a>
  - <a href="#part11">Preprocessing and data cleaning</a>
  - <a href="#part12">Visualisation using ggplot2</a>
* <a href="#part2">Part 2: MTCars and Midwest dataset (from class)</a>

***

<h2 id="part1">PART 1: Survey dataset</h2>

<center><h4 id="part11">Preprocessing and data cleaning</h4></center>

##### Import required package and dataset:
```{r}
#Packages
library(ggplot2)
library(MASS)
library(dplyr)

# Dataset from the MASS package -> survey
# I chose this because visualising the survey based on many attributes provides more options and flexibility to explore ggplot2 package. 
data("survey")
data <- survey
head(data)
```


##### Exploring the dataset
Structure:
```{r}
str(data)
```

Dimension:
```{r}
dim(data)
```

Columns or attributes:
```{r}
colnames(data)
```

##### Data cleaning
```{r}
print(paste("Number of missing values = ", sum(is.na(data))))
```
<strong>NOTE:</strong> There are many missing values that will be filled up using data imputation techniques of mean, median and mode.

##### Missing data imputation
```{r}
print(paste("Mean value of the column Wr.Hnd = ", mean(data$Wr.Hnd,na.rm = TRUE)))
print(paste("Median value of Wr.Hnd = ", median(data$Wr.Hnd,na.rm = TRUE)))
print(paste("Maximum value of Wr.Hnd = ", max(data$Wr.Hnd,na.rm = TRUE)))
print(paste("Minimum value of Wr.Hnd = ", min(data$Wr.Hnd,na.rm = TRUE)))
```
The mean and median values are very close while min and max values are having a wider difference in magnitude. Hence, the use of mean or median in substituting for a NA value in this column is a good practice. After repeating this for other columns, it was found that mean/median imputation is the best choice for other columns too (when the datatype is int or numberic). I chose Mode for categorical type attributes.


<strong>Replace numerical missig values:</strong>
```{r}
# used median value to replace the missing values
data <- data %>%
  mutate(Wr.Hnd = replace(Wr.Hnd, is.na(Wr.Hnd), median(Wr.Hnd, na.rm = TRUE)))
data <- data %>%
  mutate(NW.Hnd = replace(NW.Hnd, is.na(NW.Hnd), median(NW.Hnd, na.rm = TRUE)))
data <- data %>%
  mutate(Pulse = replace(Pulse, is.na(Pulse), median(Pulse, na.rm = TRUE)))

# used mean value to replace the missing values
data <- data %>%
  mutate(Height = replace(Height, is.na(Height), mean(Height, na.rm = TRUE)))
data <- data %>%
  mutate(Age = replace(Age, is.na(Age), mean(Age, na.rm = TRUE)))
```

<strong>Replace categorical missing values:</strong>
```{r}
# Build a function that returns the mode of the column:
get_mode <- function(x){
  distinct_values <- unique(x)
  distinct_tabulate <- tabulate(match(x, distinct_values))
  distinct_values[which.max(distinct_tabulate)]
}
# Used mode values to replace the NA values
data <- data %>% 
  mutate(M.I = if_else(is.na(M.I), get_mode(M.I), M.I))
data <- data %>% 
  mutate(W.Hnd = if_else(is.na(W.Hnd), get_mode(W.Hnd), W.Hnd))
data <- data %>% 
  mutate(Clap = if_else(is.na(Clap), get_mode(Clap), Clap))
data <- data %>% 
  mutate(Smoke = if_else(is.na(Smoke), get_mode(Smoke), Smoke))
data <- data %>% 
  mutate(Exer = if_else(is.na(Exer), get_mode(Exer), Exer))
data <- data %>% 
  mutate(Sex = if_else(is.na(Sex), get_mode(Sex), Sex))
```

##### Cleaned data ready for visualisation:
```{r}
print(paste("Number of missing values = ", sum(is.na(data))))
```

***

<center><h4 id="part12">Visualisation using ggplot2</h4></center

***

##### Scatterplot
```{r}
# Scatterplot
gg <- ggplot(data, aes(x=Height, y=Age)) + 
  geom_point(aes(col=Sex, size=Pulse)) + 
  geom_smooth(method="loess", se=F) + 
  xlim(c(130, 200)) + 
  ylim(c(0, 100)) + 
  labs(subtitle="Height Vs Age wrt Sex and sixe based on Pulse rate", 
       y="Age", 
       x="Height", 
       title="Scatterplot", 
       caption = "Source: survey")
plot(gg)
```

##### Scatterplot + encircle
```{r}
library(ggalt)
data_select <- data[data$Age > 18 & 
                            data$Age <= 20 & 
                            data$Height > 150 & 
                            data$Height < 180, ]
gg1<-ggplot(data, aes(x=Height, y=Age)) + 
  geom_point(aes(col=Sex, size=Pulse)) +   # draw points
  geom_smooth(method="loess", se=F) + 
  geom_encircle(aes(x=Height, y=Age), 
                data=data_select, 
                color="red", 
                size=2, 
                expand=0.08) +   # encircle
  labs(subtitle="Height Vs Age", 
       y="Age", 
       x="Height", 
       title="Scatterplot + Encircle", 
       caption="Source: data")
plot(gg1)
```

##### Geom_smooth line
```{r}
g <- ggplot(data, aes(Wr.Hnd, NW.Hnd))
g + geom_point() + 
  geom_smooth(method="lm", se=F) +
    labs(subtitle="Wr.Hnd vs NW.Hnd", 
       y="Wr.Hnd", 
       x="NW.Hnd", 
       title="Scatterplot with overlapping points", 
       caption="Source: data")
```

##### Jittered points
```{r}
g <- ggplot(data, aes(Wr.Hnd, NW.Hnd))
g + geom_jitter(width = .5, size=1) +
  labs(subtitle="Survey data: Wr.Hnd vs NW.Hnd", 
       y="Wr.Hnd", 
       x="NW.Hnd", 
       title="Jittered Points")
```

##### Scatterplot (counts plot)
```{r}
# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(data, aes(Wr.Hnd, NW.Hnd))
g + geom_count(col="tomato3", show.legend=F) +
  labs(subtitle="Survey data: Wr.Hnd vs NW.Hnd", 
       y="Wr.Hnd", 
       x="NW.Hnd", 
       title="Counts Plot")
```

##### Bubble Plot
```{r}
# Bubble Plot
data_select <- data[data$Smoke %in% c("Occas", "Regul"), ]
g <- ggplot(data_select, aes(Height, Age)) + 
  labs(subtitle="Survey data: Height vs Age",
       title="Bubble chart")

g + geom_jitter(aes(col=Smoke, size=Wr.Hnd)) + 
  geom_smooth(aes(col=Smoke), method="lm", se=F)
```

##### Marginal Histogram / Boxplot
```{r}
# Marginal Histogram / Boxplot
qplot(data$Pulse, geom="histogram") 
ggplot(data=data, aes(Pulse)) + 
  geom_histogram()
qplot(data$Pulse,
      geom="histogram",
      binwidth = 1,
      main = "Histogram for Pulse", 
      xlab = "Age",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.4))
ggplot(data=data, aes(Pulse)) + 
  geom_histogram(breaks=seq(10, 50, by=2), 
                 col="red", 
                 aes(fill=..count..)) +
  scale_fill_gradient("Count", low="green", high="red")+labs(title="Histogram for data", x="Pulse", y="Count")
ggplot(data=data, aes(Pulse)) + 
  geom_histogram(aes(y =..density..), 
                 breaks=seq(10, 50, by = 2), 
                 col="red", 
                 fill="green", 
                 alpha=.2) + 
  geom_density(col=2)
```

##### Stacked Bar
```{r}
#Stacked Bar
g <- ggplot(data, aes(Age)) + scale_fill_brewer(palette = "Spectral")

g + geom_histogram(aes(fill=Smoke), 
                   binwidth = .1, 
                   col="black", 
                   size=.1) +  # change binwidth
  labs(title="Histogram with Auto Binning", 
       subtitle="Smokers across Age")  

g + geom_histogram(aes(fill=Smoke), 
                   bins=5, 
                   col="black", 
                   size=.1) +   # change number of bins
  labs(title="Histogram with Fixed Bins", 
       subtitle="Smokers across Age") 

g <- ggplot(data, aes(Clap))
g + geom_bar(aes(fill=Smoke), width = 0.5) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Histogram on Categorical Variable", 
       subtitle="Smokers across various type of clappers") 
```

##### Density plot
```{r}
# Density plot
g <- ggplot(data, aes(Pulse))
g + geom_density(aes(fill=factor(Sex)), alpha=0.8) + 
  labs(title="Density plot", 
       subtitle="Pulse grouped by Sex",
       caption="Source: survey data",
       x="Pulse",
       fill="Sex")
```

##### Box Plot
```{r}
# Box Plot
g <- ggplot(data, aes(Smoke, Pulse))
g + geom_boxplot(varwidth=T, fill="plum") + 
  labs(title="Box plot", 
       subtitle="Pulse grouped by Smokers",
       caption="Source: survey data",
       x="Smokers",
       y="Pulse")
```

##### Correlation
```{r}
# correlation plot
library(ggcorrplot)

# Correlation matrix
corr <- round(cor(data[,c(2,3,6,10,12)]), 1)
corr
```

##### Plot
```{r}
# Plot
ggcorrplot(corr)
ggcorrplot(corr, method = "circle")
ggcorrplot(corr, hc.order = TRUE, outline.col = "white")
# Types of correlogram layout
# --------------------------------
# Get the lower triangle
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white")
# Get the upeper triangle
ggcorrplot(corr, hc.order = TRUE, type = "upper",
           outline.col = "white")
# Change colors and theme
# --------------------------------
# Argument colors
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white",
           ggtheme = ggplot2::theme_gray,
           colors = c("#6D9EC1", "white", "#E46726"))

ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of mtcars", 
           ggtheme=theme_bw)
```

##### Diverging Barcharts
```{r}
#Diverging bars
data1 = data[c(1:50), ]
data1$Sex <- rownames(data1)  # create new column for car names
data1$Height_z <- round((data1$Height - mean(data1$Height))/sd(data1$Height), 2)  # compute normalized mpg
data1$Height_type <- ifelse(data1$Height_z < 0, "below", "above")  # above / below avg flag
data1 <- data1[order(data1$Height_z), ]  # sort
data1$Sex <- factor(data1$Sex, levels = data1$Sex)  # convert to factor to retain sorted order in plot.

# Diverging Barcharts
ggplot(data1, aes(x=Sex, y=Height_z, label=Height_z)) + 
  geom_bar(stat='identity', aes(fill=Height_type), width=.5)  +
  scale_fill_manual(name="Height", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(subtitle="Normalised Height from 'Survey'", 
       title= "Diverging Bars") + 
  coord_flip()  
```

##### Diverging Lollipop Chart
```{r}
# Diverging Lollipop Chart
ggplot(data1, aes(x=Sex, y=Height_z, label=Height_z)) + 
  geom_point(stat='identity', fill="black", size=6)  +
  geom_segment(aes(y = 0, 
                   x = Sex, 
                   yend = Height_z, 
                   xend = Sex), 
               color = "black") +
  geom_text(color="white", size=2) +
  labs(title="Diverging Lollipop Chart", 
       subtitle="Normalized height from 'survey data': Lollipop") + 
  ylim(-2.5, 2.5) +
  coord_flip()
```

##### Diverging Dot Plot
```{r}
#Diverging Dot Plot
ggplot(data1, aes(x=Sex, y=Height_z, label=Height_z)) + 
  geom_point(stat='identity', aes(col=Height_type), size=6)  +
  scale_color_manual(name="Height", 
                     labels = c("Above Average", "Below Average"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", 
       subtitle="Normalized height from 'survey data': Dotplot") + 
  ylim(-2.5, 2.5) +
  coord_flip()

```

##### Pie chart
```{r}
# pie chart
pie <- ggplot(data, aes(x = "", fill = factor(Smoke))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="Smoke", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of Smokers", 
       caption="Source: survey data")

pie + coord_polar(theta = "y", start=0)

```

##### Violin Plot
```{r}
# Violin Plot
g <- ggplot(data, aes(Smoke, Pulse))
g + geom_violin() + 
  labs(title="Violin plot", 
       subtitle="Smokers vs Pulse",
       caption="Source: survey data",
       x="Smokers",
       y="Pulse")
```

##### Boxplot
```{r}
# Boxplot
g <- ggplot(data, aes(Smoke, Pulse))
g + geom_boxplot(varwidth=T, fill="plum") + 
  labs(title="Box plot", 
       subtitle="Pulse grouped by Smokers",
       caption="Source: survey data",
       x="Smokers",
       y="Pulse")
```

##### Boxplot
```{r}
# Boxplot
g <- ggplot(data, aes(Smoke, Pulse))
g + geom_boxplot(aes(fill=factor(Sex))) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Box plot", 
       subtitle="Pulse grouped by Smokers",
       caption="Source: survey data",
       x="Smokers",
       y="Pulse")
```

##### Dot + Box Plot
```{r}
#Dot + Box Plot
g <- ggplot(data, aes(Smoke, Pulse))
g + geom_boxplot() + 
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Box plot + Dot plot", 
       subtitle="Pulse vs Smokers: Each dot represents 1 row in source data",
       caption="Source: survey data",
       x="Smokers",
       y="Pulse")
```

##### Treemap
```{r}
library(treemapify)

ggplot(data, aes(area=Pulse, fill=Smoke, subgroup=Smoke)) + 
  geom_treemap()
ggplot(data, aes(area=Pulse, fill=Smoke, subgroup=Smoke)) + 
  geom_treemap()+
  #main group bordering
  geom_treemap_subgroup_border()+
  #subgroup heading in white
  geom_treemap_subgroup_text(color="white")+
  #all other group text in black
  geom_treemap_text(aes(label=Clap), color="black")+
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_brewer(palette = "Dark2")
```


***

<h2 id="part2">PART 2: MTCars and Midwest dataset (from class)</h2>

***

##### Import libraries and dataset
```{r}
library(ggplot2)
theme_set(theme_bw())  # pre-set the bw theme.
data("midwest")
midwest
```

##### Scatterplot
```{r}
# Scatterplot
gg <- ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) + 
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 0.1)) + 
  ylim(c(0, 500000)) + 
  labs(subtitle="Area Vs Population", 
       y="Population", 
       x="Area", 
       title="Scatterplot", 
       caption = "Source: midwest")
plot(gg)
```

##### Scatterplot + Encircle
```{r}
library(ggalt)
midwest_select <- midwest[midwest$poptotal > 350000 & 
                            midwest$poptotal <= 500000 & 
                            midwest$area > 0.01 & 
                            midwest$area < 0.1, ]
gg1<-ggplot(midwest, aes(x=area, y=poptotal)) + 
  geom_point(aes(col=state, size=popdensity)) +   # draw points
  geom_smooth(method="loess", se=F) + 
  xlim(c(0, 0.1)) + 
  ylim(c(0, 500000)) +   # draw smoothing line
  geom_encircle(aes(x=area, y=poptotal), 
                data=midwest_select, 
                color="red", 
                size=2, 
                expand=0.08) +   # encircle
  labs(subtitle="Area Vs Population", 
       y="Population", 
       x="Area", 
       title="Scatterplot + Encircle", 
       caption="Source: midwest")
plot(gg1)
```

##### Scatterplot with overlapping points
```{r}
g <- ggplot(midwest, aes(popwhite, popblack))
g + geom_point() + 
  geom_smooth(method="lm", se=F) +
    labs(subtitle="White  vs Black population", 
       y="popwhite", 
       x="popblack", 
       title="Scatterplot with overlapping points", 
       caption="Source: midwest")
```

##### Jittered Points
```{r}
data(mpg, package="ggplot2")
g <- ggplot(mpg, aes(cty, hwy))
g + geom_jitter(width = .5, size=1) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Jittered Points")
```

##### Scatterplot
```{r}
# Scatterplot
theme_set(theme_bw())  # pre-set the bw theme.
g <- ggplot(mpg, aes(cty, hwy))
g + geom_count(col="tomato3", show.legend=F) +
  labs(subtitle="mpg: city vs highway mileage", 
       y="hwy", 
       x="cty", 
       title="Counts Plot")
```

##### Bubble Plot
```{r}
# Bubble Plot
mpg_select <- mpg[mpg$manufacturer %in% c("audi", "ford", "honda", "hyundai"), ]
g <- ggplot(mpg_select, aes(displ, cty)) + 
  labs(subtitle="mpg: Displacement vs City Mileage",
       title="Bubble chart")

g + geom_jitter(aes(col=manufacturer, size=hwy)) + 
  geom_smooth(aes(col=manufacturer), method="lm", se=F)
# Marginal Histogram / Boxplot
data(mpg)
qplot(mpg$hwy, geom="histogram") 
ggplot(data=mpg, aes(mpg$hwy)) + 
  geom_histogram()
qplot(mpg$hwy,
      geom="histogram",
      binwidth = 5,  
      main = "Histogram for hwy", 
      xlab = "Age",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.4),
      xlim=c(10,50))
ggplot(data=mpg, aes(mpg$hwy)) + 
  geom_histogram(breaks=seq(10, 50, by=2), 
                 col="red", 
                 aes(fill=..count..)) +
  scale_fill_gradient("Count", low="green", high="red")+labs(title="Histogram for mpg", x="hwy", y="Count")
ggplot(data=mpg, aes(mpg$hwy)) + 
  geom_histogram(aes(y =..density..), 
                 breaks=seq(10, 50, by = 2), 
                 col="red", 
                 fill="green", 
                 alpha=.2) + 
  geom_density(col=2)
```

##### Stacked Bar
```{r}
#Stacked Bar
g <- ggplot(mpg, aes(displ)) + scale_fill_brewer(palette = "Spectral")

g + geom_histogram(aes(fill=class), 
                   binwidth = .1, 
                   col="black", 
                   size=.1) +  # change binwidth
  labs(title="Histogram with Auto Binning", 
       subtitle="Engine Displacement across Vehicle Classes")  

g + geom_histogram(aes(fill=class), 
                   bins=5, 
                   col="black", 
                   size=.1) +   # change number of bins
  labs(title="Histogram with Fixed Bins", 
       subtitle="Engine Displacement across Vehicle Classes") 

g <- ggplot(mpg, aes(manufacturer))
g + geom_bar(aes(fill=class), width = 0.5) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Histogram on Categorical Variable", 
       subtitle="Manufacturer across Vehicle Classes") 
```

##### Density plot
```{r}
# Density plot
g <- ggplot(mpg, aes(cty))
g + geom_density(aes(fill=factor(cyl)), alpha=0.8) + 
  labs(title="Density plot", 
       subtitle="City Mileage Grouped by Number of cylinders",
       caption="Source: mpg",
       x="City Mileage",
       fill="# Cylinders")
# Box Plot
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(varwidth=T, fill="plum") + 
  labs(title="Box plot", 
       subtitle="City Mileage grouped by Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

##### Correlation
```{r}
# correlation plot
library(ggcorrplot)

# Correlation matrix
data(mtcars)
corr <- round(cor(mtcars), 1)
```

##### Plot
```{r}
# Plot
ggcorrplot(corr)
ggcorrplot(corr, method = "circle")
ggcorrplot(corr, hc.order = TRUE, outline.col = "white")
# Types of correlogram layout
# --------------------------------
# Get the lower triangle
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white")
# Get the upeper triangle
ggcorrplot(corr, hc.order = TRUE, type = "upper",
           outline.col = "white")
# Change colors and theme
# --------------------------------
# Argument colors
ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white",
           ggtheme = ggplot2::theme_gray,
           colors = c("#6D9EC1", "white", "#E46726"))

ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of mtcars", 
           ggtheme=theme_bw)
```

##### Diverging Barcharts
```{r}
#Diverging bars
data("mtcars")  # load data
mtcars$`car name` <- rownames(mtcars)  # create new column for car names
mtcars$mpg_z <- round((mtcars$mpg - mean(mtcars$mpg))/sd(mtcars$mpg), 2)  # compute normalized mpg
mtcars$mpg_type <- ifelse(mtcars$mpg_z < 0, "below", "above")  # above / below avg flag
mtcars <- mtcars[order(mtcars$mpg_z), ]  # sort
mtcars$`car name` <- factor(mtcars$`car name`, levels = mtcars$`car name`)  # convert to factor to retain sorted order in plot.

# Diverging Barcharts
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) + 
  geom_bar(stat='identity', aes(fill=mpg_type), width=.5)  +
  scale_fill_manual(name="Mileage", 
                    labels = c("Above Average", "Below Average"), 
                    values = c("above"="#00ba38", "below"="#f8766d")) + 
  labs(subtitle="Normalised mileage from 'mtcars'", 
       title= "Diverging Bars") + 
  coord_flip()  
```

##### Diverging Lollipop Chart
```{r}
# Diverging Lollipop Chart
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) + 
  geom_point(stat='identity', fill="black", size=6)  +
  geom_segment(aes(y = 0, 
                   x = `car name`, 
                   yend = mpg_z, 
                   xend = `car name`), 
               color = "black") +
  geom_text(color="white", size=2) +
  labs(title="Diverging Lollipop Chart", 
       subtitle="Normalized mileage from 'mtcars': Lollipop") + 
  ylim(-2.5, 2.5) +
  coord_flip()
```

##### Diverging Dot Plot
```{r}
#Diverging Dot Plot
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) + 
  geom_point(stat='identity', aes(col=mpg_type), size=6)  +
  scale_color_manual(name="Mileage", 
                     labels = c("Above Average", "Below Average"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", 
       subtitle="Normalized mileage from 'mtcars': Dotplot") + 
  ylim(-2.5, 2.5) +
  coord_flip()

```

##### Pie chart
```{r}
# pie chart

pie <- ggplot(mpg, aes(x = "", fill = factor(class))) + 
  geom_bar(width = 1) +
  theme(axis.line = element_blank(), 
        plot.title = element_text(hjust=0.5)) + 
  labs(fill="class", 
       x=NULL, 
       y=NULL, 
       title="Pie Chart of class", 
       caption="Source: mpg")

pie + coord_polar(theta = "y", start=0)

```

##### Violin Plot
```{r}
# Violin Plot
g <- ggplot(mpg, aes(class, cty))
g + geom_violin() + 
  labs(title="Violin plot", 
       subtitle="City Mileage vs Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

##### Boxplot
```{r}
# Boxplot
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(varwidth=T, fill="plum") + 
  labs(title="Box plot", 
       subtitle="City Mileage grouped by Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

##### Boxplot
```{r}
# Boxplot
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(aes(fill=factor(cyl))) + 
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Box plot", 
       subtitle="City Mileage grouped by Class of vehicle",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

##### Dot + Box Plot
```{r}
#Dot + Box Plot
g <- ggplot(mpg, aes(manufacturer, cty))
g + geom_boxplot() + 
  geom_dotplot(binaxis='y', 
               stackdir='center', 
               dotsize = .5, 
               fill="red") +
  theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
  labs(title="Box plot + Dot plot", 
       subtitle="City Mileage vs Class: Each dot represents 1 row in source data",
       caption="Source: mpg",
       x="Class of Vehicle",
       y="City Mileage")
```

##### Treemap
```{r}
proglangs <- read.csv("https://raw.githubusercontent.com/selva86/datasets/master/proglanguages.csv")
library(treemapify)

ggplot(proglangs, aes(area=value, fill=parent, subgroup=parent)) + 
  geom_treemap()
ggplot(proglangs, aes(area=value, fill=parent, subgroup=parent)) + 
  geom_treemap()+
  #main group bordering
  geom_treemap_subgroup_border()+
  #subgroup heading in white
  geom_treemap_subgroup_text(color="white")+
  #all other group text in black
  geom_treemap_text(aes(label=id), color="black")+
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_brewer(palette = "Dark2")
```

***