---
title: 'Lab 6: igraphs'
author: "Makesh Srinivasan"
date: "10/6/2022"
output:
pdf_document: default
html_document: default
---

Registration number: 19BCE1717<br>
Faculty: Prof. Parvathi R<br>
Slot: L55 + L56<br>
Course code: CSE3020 <br>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## igraph Lab exercise

##### Do the following questions

***

##### Prerequisites
Do the following before proceeding to the questions
```{r}
# install.packages("igraph")
library(igraph) # load the library
```


##### Solutions

1.	Read the given adjacency matrix into R (adjacency.csv )
```{r}
nodes <- read.csv("Dataset2-Media-User-Example-NODES.csv") #nodes
head(nodes) 
```

***

2.	Read the given edge matrix into R(edges.csv)
```{r}
edges <- read.csv("Dataset2-Media-User-Example-EDGES.csv") #links
edges <- subset(edges, select = -c(X))
head(edges)
```

***

3.	Create and plot the graph from the adjacency matrix and edge matrix (customize the vertex color edge size, vertex frame and label)
```{r}
# The plots contain all the customisations too. But to see them individually, see the following chunk of code
net2 <- graph_from_incidence_matrix(edges)
table(V(net2)$type)
net2.bp <- bipartite.projection(net2)
as_incidence_matrix(net2)  %*% t(as_incidence_matrix(net2)) 
t(as_incidence_matrix(net2)) %*%   as_incidence_matrix(net2)
plot(net2.bp$proj1, vertex.label.color="black", vertex.label.dist=1,vertex.size=7, vertex.label=nodes$media[!is.na(nodes$media.type)])
plot(net2.bp$proj2, vertex.label.color="black", vertex.label.dist=1,vertex.size=7, vertex.label=nodes$media[ is.na(nodes$media.type)])
```

```{r}
# Individual customisations
#####edge
#colour
plot(net2.bp$proj2, edge.color="red")
#size
plot(net2.bp$proj2, edge.arrow.size=10)

#####vertex
#colour
plot(net2.bp$proj2, vertex.color="green")
#size
plot(net2.bp$proj2, vertex.size = 10)

#vector frame
plot(net2.bp$proj2, frame =TRUE, vertex.frame.color = "orange")
#label
plot(net2.bp$proj1, vertex.label.color="black", vertex.label.dist=1,vertex.size=7, vertex.label=nodes$media[!is.na(nodes$media.type)])

```

***

4.	Display the edges & vertices, the network as matrix and the names of vertices

```{r}
##### This can be printed as a matrix as shown below:
E(net2)
V(net2)
edges

##### This can also be printed as a heatmap as shown below (I used weights from a different dataset)
n <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
l <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)
l <- aggregate(l[,3], l[,-3], sum)
l <- l[order(l$from, l$to),]
colnames(l)[4] <- "weight"
rownames(l) <- NULL
net <- graph_from_data_frame(d=l, vertices=n, directed=T) 
class(net)
graph <- net
net <- simplify(net, remove.multiple = F, remove.loops = T)
as_edgelist(net, names=T)
as_adjacency_matrix(net, attr="weight")
netm <- get.adjacency(net, attr="weight", sparse=F)
colnames(netm) <- V(net)$media
rownames(netm) <- V(net)$media
palf <- colorRampPalette(c("gold", "dark orange")) 
heatmap(netm[,17:1], Rowv = NA, Colv = NA, col = palf(100), scale="none", margins=c(10,10) )
```

***

5.	Find the count of vertices and edges of the created graph
```{r}
x <- gsize(net2)
y <- gorder(net2)
print(paste("count of edges = ", x))
print(paste("count of vertices = ", y))
```

***

6.	Display the adjacency vertices of each vertex(individual) in the created graph
```{r}
adjacent_vertices(graph, V(graph), mode = "out")
adjacent_vertices(graph, V(graph), mode = "in")
adjacent_vertices(graph, V(graph), mode = "all")
adjacent_vertices(graph, V(graph), mode = "total")
```

***

7.	Find the min and max degree of the created graph
```{r}
# The degree for the paths:
degree(graph, v = V(graph), mode = c("all", "out", "in", "total"))

# Maximum degree
V(graph)$name[degree(graph)==max(degree(graph))]

#Minimum degree
V(graph)$name[degree(graph)==min(degree(graph))]
```

***

8.	Create & set vertex attribute property named profit and values("+", "-", "+", "-", "+", "-", "+", "-", "+")
```{r}
graph <- set_vertex_attr(graph,"profit",value = c("+", "-", "+", "-", "+", "-", "+", "-", "+"))
plot(graph, vertex.label.dist=1,vertex.size=7, vertex.label = c("+", "-", "+", "-", "+", "-", "+", "-", "+"))
vertex_attr(graph)
```

***

9.	Create & set vertex attribute property named type and values(either leap or non-leap year)
```{r}
graph <- set_vertex_attr(graph,"type",value = c(2000:2029))
plot(graph, vertex.label.dist=1,vertex.size=7, vertex.label = c(2000:2029))
vertex_attr(graph)
```

***

10.	Create & set edge attribute named weight and values (if edge exits in between leap year vertices then 5 else 1 )
```{r}
# The 5 are in places of nodes where the edge connects 2 leap years. Else 1
graph <- set_edge_attr(graph,"weight",index = E(graph) ,value = c("5","1","1","1","5","1","1","1","5","1","1","1","5","1","1","1","5","1","1","1","5","1","1","1","5","1","1","5","1","1","1"))
plot(graph)
get.edge.attribute(graph)
```

***

11.	Convert the created un-directed graph into directed graph based on the following rule

a. edge directed towards high value vertex
```{r}
as.directed(graph, mode = c("arbitrary"))
plot(graph)
```

b. if any one of the vertex is leap year then put the reverse edge with same weight.
```{r}
as.directed(graph, mode = c("mutual"))
plot(graph)
```

***

12. Display the adjacency matrix of the resultant directed graph.
```{r}
print(get.adjacency(graph))
```

***

13. Display the in-degree and out-degree of each vertex of resultant directed graph
```{r}
print(degree(graph,mode="in")) 
print(degree(graph,mode="out"))
```
